### 2020-10-18 api-exploration/kraken-english/Kraken_Writeup.Rmd
add english exploration

---
title: "Kraken Exploration with English Text"
author: "Yian Wang"
date: "October 17, 2020"
output:
  pdf_document: default
header-includes:
  - \usepackage[document]{ragged2e}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, echo = TRUE, fig.align="center")

```

# 1. Introduction
Kraken is an easy-to-follow system used for Optical Character Recognition (OCR) [1]. The purpose of this exploration is to look at a one-page, one-column academic sample text to ensure kraken works sufficiently for English text before training it for Japanese characters.

# 2. Data
## 2.1 Sample Text
The sample "Reconciling with History: The Chinese-Canadian Head Tax Redress" by Peter Li [2] contains varying types of text - the title, author, main text, subheading, and footnote, and watermark are all different. Thus, an area of interest will be to see how kraken interprets the different types of text, in particular, the 1 superscript in the subheading "Origin of the Chinese Head Tax in Canada". \
Keeping the green highlighted areas in the document tests the effectiveness of binarization in kraken.\
As per the tutorial, the sample is imported as a tif filetype.

```{r echo=FALSE, eval = TRUE, out.width='50%'}
knitr::include_graphics('/Users/yianwang/Desktop/kraken/inputpic.png')
```
\center __Figure 1.__ Sample Input Text: Reconciling with History

\justify
## 2.2 en_best.mlmodel
In the current kraken version to date, this model was not included but was required to run the code. It must be downloaded and imported separately (**[click here to download](https://github.com/mittagessen/kraken/issues/120)** - see mittagessen's comment on Feb 15, 2019).

# 3. Kraken Exploration

## 3.1 Local
The kraken tutorial is easily followed locally; however, for this process to be as accessible and reproducible as possible, this exploration focuses on getting the process to work on some cloud platform such as Jupyter Hub or Google Colab.

## 3.2 Jupyter Notebook/Jupyter Hub
JupyterLab on Jupyter Hub is a free cloud platform to share code.

### 3.2.1 Error
Attempting to use kraken in the JupyterLab terminal, the shell runs out of memory before the package fully installs.\ The first command 
```{bash}
pip install kraken
```
reaches a certain point in memory usage (1.84/2.00 GB) and proceeds to quit (See Appendix Figure A). The conclusion is that Jupyter Hub does not have adequate memory space so other platforms need to be explored.

## 3.3 Google Colab
Google Golab was chosen next as it has more space than JupyterHub.

### 3.3.1 Pricing
While Google Colab is free for up to 12GB of memory and run time of 12 hours, paying $9.99 a month gives access to 25GB and up to 24 hours of run time [3]. Though the free version is sufficient for the scope of this exploration task, this removes any fear of running out of space on this platform for the time being.\ Nonetheless, the discussion of pricing is important in order to gauge whether it's worth it to use the platform if/when there is a high volume of data.

### 3.3.2 Method and Outputs

First install kraken:
```{bash}
pip install kraken
```

Before beginning the process, upload the tif file of the desired text, as well as the ```en_best.mlmodel``` file to the Colab environment. \

Now, binarize the image to remove any color to allow kraken to analyze it:
```{bash}
!kraken -i image.tif bw.png binarize
```

Then, take the binarized image output and turn it into a txt file:
```{bash}
!kraken -i bw.png image.txt binarize segment ocr -m en_best.mlmodel
```
\
\newpage
Alternatively, it is possible to binarize and convert to a txt file in one line:
```{bash}
!kraken -i image.tif onestep_image.txt binarize segment ocr -m en_best.mlmodel
```

The resulting txt file is promising, with minimal errors, mostly in the title (See Appendix Figure B). \
To analyse the issue, check the line boundaries to see if kraken detected the title area (and other areas of error) correctly. A json file of the boundaries is obtained and plotted around the text to see if there exist any unexpected boundaries.\
\
Get the json file as follows: 
```{bash}
!kraken -i bw.png lines.json segment
```

Then run a function ```show_boxes``` (see Appendix 6.1) to plot these boundaries. The output is as follows:

```{r echo=FALSE, eval = TRUE, out.width='50%'}
knitr::include_graphics('/Users/yianwang/Desktop/kraken/boundaries.png')
```
\center__Figure 2.__ Sample Text with Line Boundaries

\justify
It can be see that kraken evaluated the boundaries well in this sample, with exception to the title, which has many strange boundary boxes around it. This explains the discrepancy in the produced txt file.

# 4. Evaluation

Kraken was fairly easy-to-use, and the online tutorial was very intuitive. It identified the words in the sample correctly with only minor errors. \
However, the boundaries were found at a line level, rather than a word or character level, which is less effective and sensitive. Additionally, the ```en_best.mlmodel``` file was not installed with kraken in the command line, though that issue was easily solved.

# 5. Sources
1. kraken — kraken 2.0.5-4-gbb42ba5  documentation [Internet]. Kraken.re. 2015 [cited 2020 Oct 18]. Available from: http://kraken.re/index.html \

2. Li P. Reconciling with History: The Chinese-Canadian Head Tax Redress. Journal of Chinese Overseas [Internet]. 2008 [cited 2020 Oct 18];4(1):127–40. Available from: https://brill.com/view/journals/jco/4/1/article-p127_9.xml DOI: 10.1163/179325408788691507 \

3. Google newly launches Colab Pro! - comparison of Colab and Colab pro · Buomsoo Kim [Internet]. Github.io. 2020 [cited 2020 Oct 18]. Available from: https://buomsoo-kim.github.io/colab/2020/03/15/Google-newly-launches-colab-pro.md/

# 6. Appendix
## 6.1 Code

```{python python.reticulate = FALSE}
def show_boxes(img2):
  """ 
  Find line boundaries on image text
  """
  from PIL import ImageDraw, Image
  with Image.open(img2) as img:
    drawing_object = ImageDraw.Draw(img)
    bounding_boxes = pageseg.segment(img.convert('1'), text_direction = 'horizontal-lr', 
      black_colseps = False)['boxes']
    for box in bounding_boxes:
      drawing_object.rectangle(box, fill = None, outline = 'red')
    return img
```

## 6.2 Images
```{r echo=FALSE, eval = TRUE, out.width='50%'}
knitr::include_graphics('/Users/yianwang/Desktop/kraken/memoryquit.png')
```
\center __Figure A.__ Memory in Jupyter Hub Terminal
\
\

```{r echo=FALSE, eval = TRUE, out.width='40%'}
knitr::include_graphics('/Users/yianwang/Desktop/kraken/txtfiless.png')
```
\center __Figure B.__ txt File Output