### 2020-12-08 outputs/documents/paper.Rmd
Reorganize directory and begin updating paper

---
title: "Notes and approaches to OCR"
author: "Rohan Alexander, John Tang, Diego Mamanche Castellanos"
date: "8 December 2020"
abstract: "TBD"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo=FALSE, include=FALSE, warning=FALSE}
library("reticulate")
```

# Introduction

A crucial step in the data analysis workflow is gathering data. This step involves tools and methods to collect raw data in an systematic way. In this paper we are interested in gathering data from Portable Document Format (PDF) files, in particular, when these PDFs were created by scans, and hence must be processed before their data can be analysed. Such PDF files are unstructured data, meaning that the data are not tagged and are not organised in a regular way.^[Structured data is organized in a highly regular manner or a pre-defined data model where the regularities apply to all the data in a particular dataset. Some examples are tables and relations. Semi-structured data contains this same information, but instead of having regular structures applied to all items in the dataset, the data might be interpreted with structural information. It can be supplied as tags e.g. name = "Bob" but also other markers to separate semantic elements and enforce hierarchies of records and fields within the data. Unstructured data, such as texts or images, holds information with no explicit structured data, such as tags. However, these tags may be assigned using manual or automatic techniques, converting the unstructured data to semi-structured data [@Losee2006]. This diversity and complexity of data structures complicate, even more, the data gathering process as each data type may require a specific gathering approach(es) when needed.] While tags could be assigned manually, typically a statistical model is used for reasons of efficiency, and this process is called Optical Character Recognition (OCR). We evaluate several pre-trained commercial OCR options accessed via APIs: 1) X, 2), and 3), ; additionally several OCR options that are run locally: 4) X, 5), and 6), ; and finally 7) training our own model. We find... **[TBD]**.

PDF files were developed in 1993 by the technology company Adobe and are useful for documents because they should display in a consistent way independent of the environment that created them or the environment in which they are being viewed. One feature of PDFs is that they can include a variety of objects, for instance, text, photos, figures, etc. However, this variety can limit the capacity of PDFs to be used directly as data inputs for statistical analysis. The data first needs to be extracted from the PDF.

Sometimes it is possible to copy and paste the data from the PDF. This is more likely when the PDF only contains simple text or regular tables. In particular, if the PDF has been created by an application such as Word then often the text can be extracted in this way because they are stored as text. But it is not as easy if the text has been stored as an image which is then part of the PDF

Furthermore, there is a linguistic component that adds more complexity. For instance, it is well known that non-Latin characters such as Japanese, Chinese, Arabic, among others, generate several difficulties in different stages of the text mining process **[ADD CITE]**. It is because those characters have complicated structures, a large number of categories (for instance, compare the 26 characters in the English alphabet with the thousands of characters in Japanese kanji), and resemblance among characters, font unevenness, or writing styles. In those cases, the need for reliable tools or methods with the ability to elicit data becomes extremely important.

Optical character recognition (OCR),is a process that transforms a bitmapped image of printed or handwritten text into text code **[ADD CITE]**. It thereby becomes machine-readable. It has been used to parse images of characters since the 1950s **[ADD CITE]**. These initial approaches used mechanical and optical means of rotating disks and photomultiplier flying spot scanner with a cathode ray tube lens, followed by photocells and arrays of them **[ADD CITE]**. The OCR process was slow, and one line of characters could be digitized at a time. These days once a printed or handwritten text has been captured optically by a scanner or some other optical means, the digital image goes through the following stages of a computer recognition system [@Cheriet2007]:

- The preprocessing stage that enhances the quality of the input image and locates
the data of interest.
- The feature extraction stage that captures the distinctive characteristics of the
digitized characters for recognition.
- The classification stage that processes the feature vectors to identify the characters and words.

In this paper we review various OCR options for researchers who need to gather data from PDFs in which the text must be extracted to be reused, edited, or reformatted; the text should be available for full-text information retrieval; the text is to be coded in HTML or SGML; the text should be available to adaptive equipment for the visually impaired; the file size is of concern (in terms of storage or bandwidth to transmit); or the resources are available to perform OCR and correct the output.

Although this step remains the same regardless of the field of study, the methods used vary between those fields of study because of their different interests. Moreover, the diversity of data types adds complexity to the process of gathering data.

**When they consider a primer on what OCR options exist these days depending on your technical ability and then how they perform for an easy example and then how they perform for a hard example (the Japanese extract) and finally some suggestions for what is needed in the future....**

**Our paper represents...**

**The remainder of this paper is structured as follows...**

# Data

## Test Samples

Japanese dictionaries...[Rohan]

For the purpose of this study, one page from the letter a in the Japanese dictionary was taken to assess the selected OCR tools. The image contains 3293 characters. Words in English were counted as one character for this analysis. MOreover, the sample text was divided into four images called **sections** as illustrated in Figure \@ref(fig:sections). These new images were assessed across all OCR tools.

```{r sections, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/sections.png"))
```

This example is in a vertical form, meaning that it should be read it from right to left and from the top to the bottom. This is important since it is expected to obtain from each OCR tool the correct sequence of characters. For instance, from section 1 in figure xx, the first five characters from the first line are **[一本の傘を]**. We expect to see the same sequence from each OCR tool.

The following figures corresponds to each section extracted from the sample:

```{r section1, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("inputs/cropped_pages/00-sample_pages_3_sec_1.png"))
```

```{r section2, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("inputs/cropped_pages/00-sample_pages_3_sec_2.png"))
```

```{r section3, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("inputs/cropped_pages/00-sample_pages_3_sec_3.png"))
```

```{r section4, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("inputs/cropped_pages/00-sample_pages_3_sec_4.png"))
```

## OCR options

The following list contains the OCR APIs and libraries available that will be evaluated in this study:

| OCR Tool                             | Company	 | Type           |	Monthly Price Aug/2020 |	Tech Requierements |
| ------------------------------------ | --------- | -------------- | ------- | ------------------ |
| Azure Cognitive Services (OCR)       | Microsoft | API		        | Free Web Container: 5000/Free; <br/> <br/> S1 Web Container: 0-1M/$1.50 per 1,000 trans, 1M-5M/$1 per 1,000 trans, 5M-10M/$0.65 per 1,000 trans, 10M-100M/$0.65 per 1,000 trans, +100M/$0.65 per 1,000 trans        |                    |
| Amazon Rekognition	                 | Amazon	   | API		        | First 1 million images: $0.001 per image <br/> <br/> Next 9 million images $0.0008 per image <br/> <br/> Next 90 million images $0.0006 per image <br/> <br/> Over 100 million images $0.0004 per image         |                    |
| PyTesseract (Wrapper Google's OCR Engine)    |           | Python Library | Free    |                    |
Standard: $0.002 per trans      |                    |
| Kraken	(Linux OS)                   |           | Python Library	| Free	  |                    |
| Google Vision (for Text recognition) | Google	   | API		        | 1000/Free, 5 mill/$1.5, >5mill/$0.6        |                    |
| Amazon Textract	                     | Amazon	   | API		        | First 1 Million pages: $0.015 - $0.05 per page, <br/> <br/> Over 1 Million pages: $0.01 - $0.04 per page        |                    |
|	Kindai-OCR                           | Result of n2i Project    | Python Library | Free    |                    |

## Azure Corgnitive Services - Computer Vision OCR

Among the Microsoft Azure's portfolio, Azure Cognitive Services offers Computer Vision, an API that processes images and returns information based on the visual features the user is interested in. Those services comprise object detection of an image, visual features tagging, image categorization, Optical Character Recognition (OCR), among $others^4$. The OCR API processes an image and returns the language of the document, orientation, regions, lines, boundaries, and finally, the characters.

In python we need several libraries for calling the API.

```{python eval=FALSE, include=TRUE}
import os
import sys
import requests
# If you are using a Jupyter notebook, uncomment the following line.
#%matplotlib inline
import matplotlib.pyplot as plt
from matplotlib.patches import Rectangle
from PIL import Image
from io import BytesIO
import matplotlib
from os import path
import ast
```

Then, we need to create headers and parameters necessary for calling the API. The library **requests** is commonly used in python to do that. The method **post** is needed as the API uses the HTTP method POST.

```{python eval=FALSE, include=TRUE}
#Create the URL for the API call. The endpoint corresponds to the Azure's 
# endpoint of the account.
ocr_url = endpoint + "vision/v3.0/ocr"

#Create the header using the Azure's subscription key.
headers = {'Ocp-Apim-Subscription-Key': subscription_key}

#In params language': 'ja' for japanese only. 'unk' means self detection 
params = {'language': 'unk', 'detectOrientation': 'true'} 
data = {'url': image_url}

#Call and save the response using the method post from the library request
response = requests.post(ocr_url, headers=headers, params=params, json=data)

```

Here we have an example of the response:

```{} 
{'language': 'ja',
 'textAngle': 0.0,
 'orientation': 'Up',
 'regions': [{'boundingBox': '31,30,1521,2721',
   'lines': [{'boundingBox': '730,30,44,2718',
     'words': [{'boundingBox': '735,30,34,34', 'text': 'あ'},
      {'boundingBox': '736,71,34,26', 'text': 'い'},
      {'boundingBox': '749,100,6,11', 'text': '・'},
      {'boundingBox': '735,116,34,27', 'text': 'い'},
      {'boundingBox': '736,149,33,33', 'text': 'ん'},
```

### Language and Orientation

Figure xx presents a sample that corresponds to the section 1 of the image containing Japanese text, but it also includes English text in some fragments of the text. The API offers the possibility to do self-detection, which in this case is helpful, but also the opportunity to specify the language in advance. Moreover, the orientation can also be self-detected or not.

```{r azuresection1, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/section1.png"))
```

### Regions

The blue line figure xx denotes the segment from the image that includes the characters. In this example, there is only one blue line because the document is only text. In other cases, it may contain images requiring more than one region.

```{r azureregionsection1, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/region-section1.png"))
```

### Lines

Once the region is defined, the API provides in figure xx (red rectangles) the subsets of characters of the region called Lines. From the example, we can see that some characters are not included in any of the lines. Those characters are not recognized by the API.

```{r azureregionsection2, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/lines-section1.png"))
```

### The Character Boundaries

After the lines are identified, the API provides in figure xx each recognized character bounding boxes (yellow regions). Those without a yellow square were not recognized by the API in this example.

```{r azureregionsection3, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/boundaries-section1.png"))
```

### Characters

Once the bounding boxes are located, each character takes its corresponding place, as it is shown in figure xx.

```{r azureregionsection4, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/characters-section1.png"))
```

### Advantages and Disadvantages

#### Advantages

The API is easy to use after reading the documentation. It offers an example of how to use it in different programming languages. It also offers the possibility to self-detect languages and the orientation of the document. The response is clear, and extract the information from the JSON file is not complicated.

#### Disadvantages

Even though the API offers the possibility to self-detect the language, It changes the order in which the characters should be in the output. In figure xx we can see that the first line provided by the self-detection approach is located in the middle. This situation does not allow a correct interpretation of the text because it does not follow the reading pattern. In this case, it is a vertical orientation, which means we have to read it from right to left.

```{r azureregionsection5, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/self-detection.png"))
```

When the parameter language is stated as Japanese since the beginning, as we can observe in figure xx, it recognizes the first line correctly.

```{r azureregionsection6, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/azure-images/japanese-parameter.png"))
```

## Google Vision OCR

Google Vision API can integrate vision detection features, including image labeling, face and landmark detection, optical character recognition (OCR), and tagging of explicit content. The text recognition process detects text in images and recognizes the text contained therein. Once detected, the recognizer then determines the actual text in each block and segments it into lines and words.

To evaluate this OCR option, we need the following python libraries.

```{python eval=FALSE, include=TRUE}
from os import path
from PIL import Image
from PIL import ImageFont
from PIL import ImageDraw
from PIL import ImageEnhance
import base64 #Convert an image into base64 form for API requests
import requests #API request - GET/POST
import json
import ast
```

Same as the Azure API, we need to create the headers and parameters necessary for calling the API. We also need the python library **requests** to make the API call. The method **post** is needed as the API uses the HTTP method POST.

The following example illustrates how to state the headers and parameter, but also create two functions. The first one, called encode_image converts an image into base64 format, an encoding process designed to carry data stored in binary formats across channels that only reliably support text content. It can embed image files or other binary assets inside textual assets. The second one makes an HTTP POST request that calls Google's Vision API.

```{python eval=FALSE, include=TRUE}
#Create the URL for the API call. The endpoint corresponds to the Azure's 
# endpoint of the account.
url = "https://vision.googleapis.com/v1/images:annotate"

querystring = {"key": google_vision_api_key }
headers = {'Content-Type': "application/json"}

def encode_image(image_path):
    with open(image_path, "rb") as image_file:
        return base64.b64encode(image_file.read())

def image_request(image_path):

    payload = '{\"requests\":[{\"image\":{\"content\":\"'+
    encode_image(image_path).decode('utf-8')+
    '"},\"features\":[{\"type\":\"TEXT_DETECTION\"}]}]}'

    response = requests.request("POST", url, data=payload, headers=headers, params=querystring)

    return ast.literal_eval(response.text)
```

This is an example of the response:

```{} 
{'responses': [{'textAnnotations': [{'locale': 'ja',
     'description': '-]\n一\n一本の傘を二人でさすこと。多く男女二人の場合に一あい-えん [哀婉](名) 哀れにやさしいこと。 こと。しん【愛郷心](名) 愛婚の精神。.............名)あいく\n-[製~Q)
     れ]()あ\nの。0子持鮎の願漬。\nあいきょうp®(愛郷](名)生まれ故郷を愛する\nに\nVド\nし
     いこと。叉、その程度。\n信的思想た打破すること。\n',
     'boundingPoly': {'vertices': [{'x': 26, 'y': 21},
       {'x': 1556, 'y': 21},
       {'x': 1556, 'y': 2748},
       {'x': 26, 'y': 2748}]}},
    {'description': '-',
     'boundingPoly': {'vertices': [{'x': 1227, 'y': 686},
       {'x': 1227, 'y': 677},
       {'x': 1262, 'y': 677},
       {'x': 1262, 'y': 686}]}},

       すること。\n'}}]}
           {'description': '誘',
     'boundingPoly': {'vertices': [{'x': 785, 'y': 880},
       {'x': 814, 'y': 880},
       {'x': 814, 'y': 914},
       {'x': 785, 'y': 914}]}},
    {'description': '(',
     'boundingPoly': {'vertices': [{'x': 775, 'y': 916},
       {'x': 814, 'y': 916},
       {'x': 814, 'y': 940},
       {'x': 775, 'y': 940}]}},
    ...],
```

From the previous example, we can observe that the response returns a list of nested dictionaries. The first nested dictionary called **textAnnotations** consists of a list of dictionaries. Each one of those has a description (character) and boundingPoly (boundaries of the characters).

### BoundingPoly

Each boundingPoly consist of vertices in which the symbol is located.

#### Region

Te first object in the nested dictionary contains all identified characters in the image as well as the boundaries that comprises the location of those characters. This object can resemble the idea of the region provided in the Azure's API.

Figure xx shows the region identified by the API.

```{r azureregionsection7, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/google-images/sections.png"))
```

#### The Character Boundaries

The remaining objects enclose the boundaries of the characters. As it is shown in figure xx, each boundingPoly node can include one or more characters.

```{r azureregionsection8, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/google-images/boundaries.png"))
```

### Characters

Once the bounding boxes are located, each character takes its corresponding place, as it is shown in figure xx. **Pending find a way to plot the text in vertical form to make it readable.**

```{r azureregionsection9, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/google-images/characters.png"))
```

### languageHints

Google Vision API offers the advantage of including in the request language hints. It consists of a list of languages the user knows in advance the image contains. For this study, we run two versions, one without hints and the other one with a list of two hints, such as Japanese and English, as the image contains some words in English. Both versions returned the same output.

### Advantages and Disadvantages

#### Advantages

Same as the Azure API, this is also easy to use after reading the documentation. It also offers several examples in different programming languages using the Google libraries for Python. It gives the possibility to include languages in advanced as hints, and identifies the orientation automatically. The response contains an object with all identified characters in the correct order.

Moreover, the API gives the opportunity to send the image in base64 format or using an internal (google storage) or external URL.

#### Disadvantages

Unlike Azure Cognitive Services, which returns each recognized character individually, Google Vision groups characters in some cases. This situation adds some degree of complexity when processing the response from the API. **[Not sure if it is a disadvantage in real life]**

## PyTesseract - Python Library

Python-Tesseract is an optical character recognition (OCR) library for python. It is a wrapper for Google's Tesseract-OCR Engine. PyTesseract can read all image types such as png, jpeg, gif, tiff, BMP, among others. It is widely used to process everything from scanned documents. This inbuilt python library performs the OCR techniques on the image to produce the digitized output text in a readable and editable form.

To use this library, we need to install Tesseract. For this study, we use version 4. It is necessary to call the .exe file to use Tesseract with PyTesseract. See the following example.

```{python eval=FALSE, include=TRUE}
#Call .exe location for Windows
pytesseract.pytesseract.tesseract_cmd = r'C:\Program Files\Tesseract-OCR\tesseract.exe' 
```

Along with the PyTesseract library, the following libraries are used.

```{python eval=FALSE, include=TRUE}
from PIL import Image
from PIL import ImageOps
import pytesseract
from os import path
import pandas as pd
import matplotlib.pyplot as plt
from matplotlib.patches import Rectangle

```

PyTesseract offers several methods with different outputs when processing an image to extract text. According to Pypi.org, the uses of those methods are:

- **image_to_string:** Returns the result of a Tesseract OCR run on the image to string
- **image_to_boxes** Returns result containing recognized characters and their box boundaries
- **image_to_data:** Returns result containing box boundaries, confidences, and other information. Requires Tesseract 3.05+. For more information, please check the Tesseract TSV documentation
- **image_to_osd:** Returns result containing information about orientation and script detection.
- **image_to_alto_xml:** Returns result in the form of Tesseract’s ALTO XML format.
- **run_and_get_output:** Returns the raw output from Tesseract OCR. Gives a bit more control over the parameters that are sent to tesseract.

For this analysis, we use the method image_to_data as this offers the boundaries and the recognized characters in a handy form so we can understand the process. Furthermore, in PyTesseract, it is needed to specify the language in advance. Since PyTesseract is a wrapper of Google's Tesseract-OCR Engine, Tesseract-OCR must have installed all required languages. In this case, it is Japanese, which is not included in its out-of-the-box version. To include a new language, adding the **.traineddata** file into the Tesseract-OCR/tessdata folder is sufficient. However, since the text in the image is vertical Japanese. It is necessary to include two files: jpn.traineddata and jpn_vert.traineddata. Once the files are located correctly, the following code extracts the text from the image.

```{python eval=FALSE, include=TRUE}
img1 = Image.open(r'~\inputs\00-sample-pages_1.png')
image_to_data = pytesseract.image_to_data(img1, lang = 'jpn_vert')

```

The output from the previous code converted into a pandas dataframe is shown in figure xx.

```{r azureregionsectio98, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/figure11.png"))
```

From figure xx, we can observe that the output in the column **level** generates five levels. Those are pages, blocks, paragraphs, lines, and words (characters). Each one is a subset of the previous one.

### Page

For this example, the page resembles the concept of region, which is the largest area of the image enclosing all recognized characters. Figure 12 illustrates the page area.

```{r azureregionsectio12, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/pytesseract-images/page.png"))
```

### Blocks

The blocks are subsets of the page area that consist of small groups of recognized characters. In figure 13, we can see that most of the blocks are well ordered. In terms of correctness of the boundaries, we can see that some of them overlap several characters that are not recognized by the library.

```{r azureregionsectio13, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/pytesseract-images/blocks.png"))
```

### Paragraphs

Similar to the blocks, the paragraphs are subsets of blocks. These follow the same idea of subareas containing a set of characters. In this case, for example, we can look at paragraphs three and four, which are subsets of block four (figure xx). It is interesting to see how the only paragraph four of block four is a very small area compared to its parent block. This situation happens in several other blocks. See figure xx for a better understanding.

```{r azureregionsectio14, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/pytesseract-images/paragraphs.png"))
```

### Lines

Once the paragraphs are defined, the library provides in figure xx, the last subsets of characters called Lines. From the example, we can see that most of the characters are located in one of the lines. However, several others are not. For some cases there is no clear reason why those characters are not included. For instance, between the lines 14 and 15, two clear lines are not recognized by the library.

```{r azureregionsectio15, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/pytesseract-images/lines.png"))
```

### Character Boundaries

These boundaries enclose the characters. Figure xx shows that a character boundary can include one or more characters. However, some character boundaries overlap each other. Moreover, same as the previous areas, several characters have been skipped with no apparent reason.

```{r azureregionsectio16, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/pytesseract-images/boundaries.png"))
```

### Characters

In terms of character-level recognition, we can observe that many characters are not recognized. Interestingly those characters are in a correct block, paragraph, line, and character boundary, but it was skipped. For example, in the first line of recognized characters and the first character boundary of this line, there are two words, but one of them is not recognized. Figure xx illustrates this case.

```{r azureregionsectio17, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/pytesseract-images/characters.png"))
```

### Advantages and Disadvantages

#### Advantages

The most important advantage is that this is a free library. It is also a wrapper from Google's Tesseract Engine. Moreover, the way it presents the covered areas are very detailed. It offers blocks, paragraphs, lines, boundaries, and finally the character level.

#### Disadvantages

Some lines were completely ignored by Pyteseract, even thought it was clear enough to be recognized by the tool. Moreover, same as Google Vision, this library groups more than one character sometimes.

## Kindai-OCR

According to the Center for Open Data in the Humanities, Kindai-OCR is an OCR system for modern Japanese documents. It was built under the N2I project that study state-of-the-art statistical models for OCR, and work on the development of infrastructure and providing open access to data. The datasets were constructed by the University of Tokyo Center for Education and Research and the National Institute of Japanese Language and are used for OCR machine learning. These datasets are commonly known as the Modern Magazine Datasets.

The Kindai-OCR code was developed in Python and released as an open source code in August 2020 on Github. The Github folder contains all the necessary code and instructions to assess the OCR tool. **[not sure if it is necessary to mention the modification in the file test.py I did, in order to run the model using GPUs.]**

After running the Kindai-OCR for our samples, the system generated two main outputs. They are explained as follows:

### File result.xml

This file is presented in an Extensible Markup Language (XML) format containing a tree with the identified lines, as well as the corresponding characters. An example of the xml file is shown as follows:

```{}
<?xml version='1.0' encoding='Shift_JIS'?>
<paper xmlns="http://codh.rois.ac.jp/modern-magazine/"><page dpi="100" file="00-sample_pages_3_sec_3.png" height="1040" number="1" width="2430"><line height="905" width="72" x="332" y="0">あいくるし發路事、愛くるし（形、</line><line height="1026" width="76" x="466" y="0">あいくち島（合巳。名。あれせも。ものよ</line><line height="1026" width="69" x="592" y="0">あいくすりし（（きき」名その人の爲に信身あ</line><line height="1031" width="65" x="660" y="0">あいきん（及兵」（名起しい得をうつし表しした</line><line height="1031" width="78" x="778" y="0">あいきん愛心。名客する東なふする、</line><line height="361" width="72" x="849" y="0">あいきよう對</line><line height="292" width="64" x="918" y="0">あいきよう</line><line height="1031" width="64" x="981" y="0">あいきよう對（愛婦（名（（（らきささ。（（彼。。</line><line height="292" width="64" x="2264" y="0">あいきまう</line><line height="997" width="69" x="1749" y="34">（愛教商）名」名毎金。造。。愛もがなてては立</line><line height="337" width="55" x="286" y="38">府がはゆらしい。</line><line height="930" width="68" x="407" y="38">くあふ人。○つばのない兌。。ヒ直。九五五五。</line>
```

After transforming this file into a dataframe form, we found that 238 lines with their corresponding characters, were generated. In figure XX we can see that the result.xml file provides information such as the corresponding level (paper, page, or line) in the tag column, the identified characters in the text column, the associated image the file column, among other attributes.

```{r azureregionsectio27, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/kindai-images/xml-processed.png"))
```

### New version of the image

Kindai-OCR generated a new version of the image containing the identified lines and characters. An example is shown as follows in figure xx.

```{r azureregionsectio47, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/kindai-images/kindai-section1.jpg"))
```

### Advantages and Disadvantages

#### Advantages

Same as libraries such as Pytesseract, one of The most important advantages is that it is a free tool. Moreover, the recognition process is very easy. To obtain results, just by putting the images in the test folder and running the file called test.py on Python is enough to obtain the results.

#### Disadvantages

At the time of this assessment, the use of GPUs is required which is a limitation to take into account. For this study, we used Colab from Google, which offers this capability. Moreover, Kindai-OCR returns the characters grouped by lines. They are not generated individually.

## Kraken - Python Library

Kraken is a python library created by Benjamin Kiessling from Université PSL in France and Leipzig University. This library is a combination of a fork (independent development built on an existing one) from the Ocropus package and CLSTM neural network library. Its purpose was to improve accuracy and performance rates compared with Ocropus. It has been tested with languages such as Arabic, Persian, Syriac Polytonic Greek, Latin, among other. However, there is no information about japanese-trained models.

To assess the performance of this library, it is necessary to train a model for japanese texts. To do that, Kraken offers a  function called **train**. This function expects a series of images that according to Kraken´s website, these images must be high quality scans, preferably color or grayscale, and at least 300dpi. Images in PDF format are not allowed. Each image must be accompanied with a text file with extension .gt.txt. This file will contain the character(s) that correspond(s) to the image with the same name. For instance in figure xx we see the image file is called **sec1_line1.png**, the pair-wise text element for this image is **sec1_line1.gt.txt**.

```{r azureregionsectio18, echo=FALSE, fig.cap="UPDATE CAPTION", out.width = '90%'}
knitr::include_graphics(here::here("outputs/documents/figure18.png"))
```

After the training stage, Kraken analyses the layout and extracts text lines from an input image for later processing by the recognition. A step by step process is provided by Kraken´s website (http://kraken.re/index.html) to recognize text from an image using the model we want.

### Training process challenges

The training process itself is very friendly. However, the data preparation is the most important part as it generates the samples we feed into the model. Other studies pointed out that for languages such as Arabic, at least 800 lines are required to train a descent model. For the Japanese text of this study we trained 36 lines in which several characters appear more than once. The result after the training process was not good, indicating that the need of greater sample data is needed, which is not in the scope of this study.

# Other limitations

In order to use this library it is required to have either Linux or Mac. In this study, we use Google Colab which offers a Linux environment to assess the library.

## Amazon Textract

According to Amazon's website, Amazon Textract can detect Latin-script characters from the standard English alphabet and ASCII symbols.

https://aws.amazon.com/textract/faqs/

## Amazon Rekognition

According to Amazon's website, Amazon Rekognition supports text in most Latin scripts and numbers. Text detection recognizes up to 50 sequences of characters per the image or video frame and lists them as words and lines.

https://aws.amazon.com/rekognition/faqs/

# Evaluation

For each option, we rank them based on two scales: results and difficulty. We need to work out a definition for each of these.

Diego's comment: From what I learned so far, it could be easiness (an API is more straightforward than a library), number of characters correctly recognized, flexibility, and cost. Moreover, some APIs such as Google's groups more than one character, whereas Azure returns only one character per boundary.

In terms of flexibility...

Sequence of the characters...

Output given...xml, json, flat file, csv...

Number of characters well recognized...(Accuracy)

The main aspect of the end product will be a graph with two axis, and dots for where each service is positioned, coloured or faceted by the test.

# Discussion

# Appendix

# References

1: https://www.researchgate.net/publication/267465115_An_Overview_and_Applications_of_Optical_Character_Recognition

2: https://books-scholarsportal-info.myaccess.library.utoronto.ca/en/read?id=/ebooks/ebooks2/wiley/2011-12-13/1/9780470176535

3:

4: https://docs.microsoft.com/en-us/azure/cognitive-services/computer-vision/home

5: https://docs.microsoft.com/en-us/azure/cognitive-services/computer-vision/quickstarts/python-print-text

6: https://cloud.google.com/vision/docs/ocr

7: https://pypi.org/project/pytesseract/

8: https://github.com/tesseract-ocr/tessdata

9: https://github.com/UB-Mannheim/tesseract/wiki

10: http://codh.rois.ac.jp/software/kindai-ocr/

11: https://github.com/ducanh841988/Kindai-OCR

12: Anh Duc Le, Daichi Mochihashi, Katsuya Masuda, Hideki Mima, and Nam Tuan Ly. 2019. Recognition of Japanese historical text lines by an attention-based encoder-decoder and text line generation. In Proceedings of the 5th International Workshop on Historical Document Imaging and Processing (HIP ’19). Association for Computing Machinery, New York, NY, USA, 37–41. DOI:https://doi.org/10.1145/3352631.3352641